Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-19 paper/sheaf_lics.tex

In this paper, we defined a way to extend type theory with new
principles, using the already known method in topos theory,
sheafification. Using this method, one an add some principles, like
excluded middle, to type theory, but with computational content, while
doing it with axiom break the extraction of proofs.

Beyond the result, this work allows to illustrate some homotopy type theoritical
techniques, as well as use of proof assistants.

\kq{J'ai un peu copié sur la conclusion de Licata ``Eilenberg MacLane
  spaces''. Si ça pose problème, je changerai.}

In future works, we would like to improve this construction in two
ways:
\begin{itemize}
\item At the moment, the sheafification process only works with
  truncated types. But some classical types in homotopy type theory
  are not truncated, like the circle $\mathbb S^1$. 
  Thus, we plan to extend the sheafification to all types, even if non
  truncated
\item In this work, the Giraud-Rezk-Lurie axiom is admitted. The work
  done in~\cite{lurie} suggests, if we see homotopy type theory as
  higher topos theory, that this axiom could be provable. A
  computer-checked proof of this fact would improve our developpement.
\end{itemize}
Moreover, in topos theory, we know that Lawvere-Tierney subsumes
Grothendieck (\cite{maclanemoerdijk}, section~V.4). Higher Lawvere-Tierney sheaves are
presented here, and higher Grothendieck sheaves in~\cite{lurie} ; we want to
check if this remains true in higher topos theory.

@@

We would like to thank all maintainers of the Coq/HoTT library, who
allowed us to base our formalization on something. We would also like
to thank Bas Spitters for the helpful discussions on the subject.

>>>>

In this paper, we have defined a way to extend homotopy type theory
with new principles, leveraging Lawvere-Tierney sheafification to
higher topos theory. 
%
Using sheafification, it is possible to add new logical principles,
like the law of excluded middle, to homotopy type theory and this
without axiom.
%
Beyond the result, our work is part of a more general line of works
which aim to illustrate that homotopy type theory is a very promising
theory for formalizing mathematics inside a proof assistant.

In future works, we would like to improve this construction in two
ways:
\begin{itemize}
\item 
  At the moment, sheafification only works with truncated
  types. Also this is fine from a programming point of view, 
  many important types in homotopy theory are not
  truncated, like the circle $\mathbb S^1$.  Thus, we plan to extend
  the sheafification to all types, even if non truncated by using a
  limit construction.
\item 
  In this work, a particular case of a Giraud-Rezk-Lurie axiom has
  been admitted. The work done in~\cite[Chapter 6]{lurie} suggests,
  considering that homotopy type theory is a language for higher topos
  theory, that this axiom is provable. We plan to investigate if the
  higher topos theoretic proof can be adapted to our setting. 
  %
  % A computer-checked proof of this fact would improve our
  % developpement.
\end{itemize}
Moreover, in topos theory, Lawvere-Tierney subsumes Grothendieck
(\cite{maclanemoerdijk}, section~V.4) in the sense that any
Grothendieck topology gives rise to a Lawvere-Tierney topology with
the same notion of sheaves. Higher Lawvere-Tierney sheaves are
presented here, and higher Grothendieck sheaves have been defined
in~\cite{lurie}. It should be interesting to check if the subsumption
remains true in higher topos theory.

@@

We would like to thank all maintainers of the Coq/HoTT library, who
provide a very solid base for our formalization. We also would like
to thank Bas Spitters for helpful discussions on the subject.
